---
title: "tree plot combined_5th OCT 2019"
output:
  word_document: default
  pdf_document:
    fig_height: 6
    fig_width: 8
  html_document: default
editor_options:
  chunk_output_type: inline
---

#### IMPORTANT NOTE: If you wish to have the combined figure in this doc then you have the 'resprout_failure' markdown first. Until I figure out how to save and load ggplot images easily....


```{r setup, include=TRUE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=8, fig.height=6)

library(here)
library(dplyr)
library(ggplot2)
library(mgcv)
library(readxl)
library(stringr)
library(tidyr)
library(bayesplot)
library(brms)
library(tidybayes)
library(RColorBrewer)
library(rstan)
library(splines)
library(forcats)
library(ggthemes)
library(ggpubr)
library(egg)

#These options help Stan run faster:

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

# Make sure the scales package is available (it should be if ggplot is installed)
requireNamespace("scales")

# Default graph theme - white background
theme_set( theme_bw() )

set.seed(42)


# Set to TRUE to force models to be refitted
REFIT_MODELS <- FALSE


# Create a folder for fitted models if one does not already exist
FittedModelsPath <- here("fitted_models")
if (!dir.exists(FittedModelsPath)) {
  dir.create(FittedModelsPath)
}


# Load brms if it is installed (requires Stan software and the rstan package).
# If not installed, HAS_BRMS will be set to FALSE.
HAS_BRMS <- suppressWarnings(
  require("brms", quietly = TRUE)
)


###### Some helper functions #####

# Calculate standard page sizes
pagesize <- function(size = c("A4", "A3", "A2", "A1", "A0"), 
                     orientation = c("portrait", "landscape"),
                     units = c("cm", "mm")) {
  
  size <- match.arg(size)
  orientation <- match.arg(orientation)
  units <- match.arg(units)
  
  alpha <- 1000 * 2^(1/4)
  i <- as.integer(substr(size, 2, 2))
  long <- alpha * 2^(-i/2)
  
  page <- switch(
    orientation,
    portrait = c(width = long / sqrt(2), height = long),
    landscape = c(width = long, height = long / sqrt(2))
  )
  
  page <- round(page)
  if (units == "cm") page <- page / 10
  
  page <- c(as.list(page), units = units)
  class(page) <- "pagesize"
  
  page
}



# Save a graph to a PDF file
gg_pdf <- function(plot, filename, size = pagesize("A4", "landscape", "cm")) {
  
  if (!inherits(size, "pagesize")) stop("The size argument should be a pagesize (list) object")
  
  ggsave(
    filename, 
    plot, 
    width = size$width,
    height = size$height,
    units = size$units)
}


# Calculate highest posterior density interval for a vector of values
hpdi.vec <- function (x, prob = 0.95) {
  n <- length(x)
  if (n <= 1) stop("x must have more than 1 element")
  x <- sort(x)

  gap <- max(1, min(n - 1, round(n * prob)))
  init <- 1:(n - gap)

  inds <- which.min(x[init + gap] - x[init])

  out <- c(lower = x[inds], upper = x[inds + gap])
  out
}


```

### sapling mortality 
```{r fig.height = 4, fig.width = 6.35}
library(ggpubr)
library(egg)
fig <- egg::ggarrange(DSF.saplings, WSF.saplings, ncol = 2, nrow = 1) 

combined_sap_mortality <- annotate_figure(fig, bottom = text_grob("Drought / fire combination", face = "bold", vjust = -1.5, hjust = 0.45), left = text_grob("P mortality (%)", face = "bold", rot = 90, vjust = 0.9, hjust = 0.3))  
  

ggsave("combined_sap_mortality.png")

combined_sap_mortality
```


### juvenile abundance plot


```{r fig.height = 4, fig.width = 6.35}
library(ggpubr)
library(egg)
fig <- egg::ggarrange(DSF.abund2, WSF.abund2, ncol = 2, nrow = 1) 

combined_plot_abund <- annotate_figure(fig, bottom = text_grob("Drought / fire combination", face = "bold", vjust = -1.5, hjust = 0.45), left = text_grob("Predicted count per site", face = "bold", rot = 90, vjust = 0.9, hjust = 0.3))
  

ggsave("combined_abundance.png")

combined_plot_abund
```

### combined_plot_mature_abund.replace


```{r fig.height = 3.5, fig.width = 6.35}
library(ggpubr)
library(egg)
fig <- egg::ggarrange(DSF.replace_mag_jags_mature, WSF.replace_mag_jags_mature, DSF.mature.abund, WSF.mature.abund, ncol = 4, nrow = 1) 

combined_plot_mature_abund.replace <- annotate_figure(fig, top = text_grob("Replacement                                             Abundance", face = "bold", vjust = 0.2, hjust = 0.46), bottom = text_grob("Drought / fire combination", face = "bold", vjust = -1.5, hjust = 0.45), left = text_grob("Predicted count per site", face = "bold", rot = 90, vjust = 0.9, hjust = 0.35))
  

ggsave("combined_plot_mature_abund.replace.png")

combined_plot_mature_abund.replace
```


### DSF recruitment plot ##v1


```{r fig.height = 4, fig.width = 8}
library(ggpubr)
library(egg)
fig <- egg::ggarrange(DSF.nseed, DSF.ndead, DSF.replace_mag_jags, ncol = 3, nrow = 1) 

combined_DSF_plot <- annotate_figure(fig, bottom = text_grob("Drought / fire combination", face = "bold", vjust = -0.5, hjust = 0.4), left = text_grob("Predicted count per site", face = "bold", rot = 90, vjust = 0.9, hjust = 0.4)) 
  

ggsave("combined_recruit_plot_DSF.png")

combined_DSF_plot
```


## v2

```{r fig.height = 6.25, fig.width = 8}
library(ggpubr)
library(egg)
fig <- egg::ggarrange(DSF.nseed, DSF.ndead, DSF.replace_mag_jags, WSF.nseed, WSF.ndead, WSF.replace_mag_jags, ncol = 3, nrow = 2) 

combined_plot.v2 <- annotate_figure(fig, bottom = text_grob("Drought / fire combination", face = "bold", vjust = -0.5, hjust = 0.4), left = text_grob("Predicted count per site", face = "bold", rot = 90, vjust = 0.9, hjust = 0.4), right = text_grob("DSF                                                   WSF", face = "bold", rot = 270, vjust = 0.5, hjust = 0.6)) 
  

ggsave("combined_recruit_plot.v2.png")

combined_plot.v2
```

### WSF recruitment plot v1
```{r fig.height = 4, fig.width = 8}
library(ggpubr)
library(egg)
fig <- egg::ggarrange(WSF.nseed, WSF.ndead, WSF.replace_mag_jags, ncol = 3, nrow = 1) 

combined_WSF_plot <- annotate_figure(fig, bottom = text_grob("Drought / fire combination", face = "bold", vjust = -0.5, hjust = 0.4), left = text_grob("Predicted count per site", face = "bold", rot = 90, vjust = 0.9, hjust = 0.4)) 
  

ggsave("combined_recruit_plot_WSF.png")

combined_WSF_plot
```

## WSF v2

```{r fig.height = 4, fig.width = 8}
library(ggpubr)
library(egg)
fig <- egg::ggarrange(WSF.nseed, WSF.ndead, WSF.replace_mag_jags, ncol = 3, nrow = 1) 

combined_WSF_plot.v2 <- annotate_figure(fig, bottom = text_grob("Drought / fire combination", face = "bold", vjust = -0.5, hjust = 0.4), left = text_grob("Predicted count per site", face = "bold", rot = 90, vjust = 0.9, hjust = 0.4)) 
  

ggsave("combined_recruit_plot_WSF.v2.png")

combined_WSF_plot.v2
```

### mature tree replacement


```{r fig.height = 4, fig.width = 6.35}
library(ggpubr)
library(egg)
fig <- egg::ggarrange(DSF.replace_mag_jags_mature, WSF.replace_mag_jags_mature, ncol = 2, nrow = 1) 

combined_replacment_mature <- annotate_figure(fig, bottom = text_grob("Drought / fire combination", face = "bold", vjust = -1.5, hjust = 0.45), left = text_grob("Predicted count per site", face = "bold", rot = 90, vjust = 0.9, hjust = 0.3))
  

ggsave("combined_replacment_mature.png")

combined_replacment_mature
```

## WSF mortality plot
```{r fig.height = 6, fig.width = 5}
library(ggpubr)
library(egg)
fig <- egg::ggarrange(WSF.mature.DF, WSF.mature.BK, ncol = 1, nrow = 2) 

combined_WSF_plot <- annotate_figure(fig, bottom = text_grob("DBH (cm)", face = "bold", vjust = 0.8, hjust = 0.2), top = text_grob("Fire scar presence", face = "bold", hjust = 0.35, vjust = 3), left = text_grob("P mortality (%)", face = "bold", rot = 90, vjust = 0.9), right = text_grob("Drought / fire effect                             Bark type effect\n", face = "bold", rot = 270, hjust = 0.5, vjust = 0.9)) 
  

ggsave("combined_RF_WSFplot.png")
```

## DSF mortality plot

```{r fig.height = 6, fig.width = 5}
library(ggpubr)
library(egg)
fig <- egg::ggarrange(DSF.mature.DF, DSF.mature.BK, ncol = 1, nrow = 2) 

combined_DSF_plot <- annotate_figure(fig, bottom = text_grob("DBH (cm)", face = "bold", vjust = 0.8, hjust = 0.2), top = text_grob("Fire scar presence", face = "bold", hjust = 0.35, vjust = 3), left = text_grob("P mortality (%)", face = "bold", rot = 90, vjust = 0.9), right = text_grob("Drought / fire effect                           Bark type effect\n", face = "bold", rot = 270, hjust = 0.5, vjust = 0.9))

ggsave("combined_RF_DSFplot.png")
```



#```{r fig.height = 6, fig.width = 7}
fig <- egg::ggarrange(DSF.RF.poles.diff.scar, DSF.RF.mature.diff.scar, DSF.RF.poles.diff.bark, DSF.RF.mature.diff.bark, ncol = 1, nrow = 4) 

combined_DSF_RF_diff_plot <- annotate_figure(fig, bottom = text_grob("Drought severity", face = "bold", vjust = -4.7, hjust = 0.4), top = text_grob("Bark type", face = "bold", vjust = 0.4, hjust = 0.3), left = text_grob("\u0394 P fail d/t bark type (%)         \u0394 P fail d/t fire scar (%)", face = "bold", rot = 90, vjust = 0.1, hjust = 0.45), right = text_grob("                      P-S                 M                      P-S                   M\n", face = "bold", rot = 270, hjust = 0.67, vjust = 0.9))

combined_DSF_RF_diff_plot <-  annotate_figure(combined_DSF_RF_diff_plot, top = text_grob("-----------------------------------------------------------------------------------------------------------", vjust = 23.5)) 

ggsave("combined_RF_DSF_diffplot.png")

combined_DSF_RF_diff_plot
```


```{r fig.height = 6, fig.width = 7}
fig <- egg::ggarrange(WSF.RF.poles.diff.scar, WSF.RF.mature.diff.scar, WSF.RF.poles.diff.bark, WSF.RF.mature.diff.bark, ncol = 1, nrow = 4) 

combined_WSF_RF_diff_plot <- annotate_figure(fig, bottom = text_grob("Drought severity", face = "bold", vjust = -4.7, hjust = 0.4), top = text_grob("Bark type", face = "bold", vjust = 0.4, hjust = 0.3), left = text_grob("\u0394 P fail d/t bark type (%)         \u0394 P fail d/t fire scar (%)", face = "bold", rot = 90, vjust = 0.1, hjust = 0.45), right = text_grob("                      P-S                 M                      P-S                   M\n", face = "bold", rot = 270, hjust = 0.67, vjust = 0.9))

combined_WSF_RF_diff_plot <-  annotate_figure(combined_WSF_RF_diff_plot, top = text_grob("-----------------------------------------------------------------------------------------------------------", vjust = 23.5)) 

ggsave("combined_RF_WSF_diffplot.png")

combined_WSF_RF_diff_plot
```




```{r fig.height = 8, fig.width = 7}

fig <- egg::ggarrange(WSF.poles, WSF.mature, ncol = 1, nrow = 2) 

combined_WSF_plot <- annotate_figure(fig, bottom = text_grob("Fire scar presence", face = "bold", vjust = -0.5), top = text_grob("Drought / fire combination", face = "bold", vjust = 2.5), left = text_grob("P resprouting failure (%)", face = "bold", rot = 90, vjust = 0.9), right = text_grob("Pole-sized trees                                                    Mature trees\n", face = "bold", rot = 270, hjust = 0.5, vjust = 0.9))

ggsave("combined_RF_WSFplot.png")
```


```{r fig.height = 4, fig.width = 7}
fig <- egg::ggarrange(WSF.poles.diff, WSF.mature.diff, ncol = 1, nrow = 2) 

combined_DSF_diff_plot <- annotate_figure(fig, top = text_grob("Bark type", face = "bold", vjust = 0.6, hjust = 0.3), left = text_grob("\u0394 P resprouting failure (%)", face = "bold", rot = 90, vjust = 0.2, hjust = 0.4), right = text_grob("Pole-sized         Mature\n", face = "bold", rot = 270, hjust = 0.68, vjust = 0.9))

ggsave("combined_RF_WSF_diffplot.png")
```


```{r fig.height = 8, fig.width = 7}
library(ggpubr)
library(egg)
fig <- egg::ggarrange(DSF.RP.poles.diff.scar, DSF.RP.poles.diff.bark, ncol = 1, nrow = 2) 

combined_DSF_poles_diff_plot <- annotate_figure(fig, bottom = text_grob("Drought severity", face = "bold", vjust = -4.7), top = text_grob("Bark type", face = "bold", vjust = 0.4), left = text_grob("\u0394 P membership due to bark type (%)       \u0394 P membership due to fire scar (%)", face = "bold", rot = 90, vjust = 0.1, hjust = 0.47), right = text_grob("Resprouting position", face = "bold", rot = 270, hjust = 0.65, vjust = 0.25))

combined_DSF_poles_diff_plot <-  annotate_figure(combined_DSF_poles_diff_plot, top = text_grob("-----------------------------------------------------------------------------------------------------------", vjust = 31.7)) 

ggsave("combined_DSF_poles_diff_plot.png")


```

```{r fig.height = 8, fig.width = 7}
library(ggpubr)
library(egg)
fig <- egg::ggarrange(DSF.RP.mature.diff.scar, DSF.RP.mature.diff.bark, ncol = 1, nrow = 2) 

combined_DSF_mature_diff_plot <- annotate_figure(fig, bottom = text_grob("Drought severity", face = "bold", vjust = -4.7), top = text_grob("Bark type", face = "bold", vjust = 0.4), left = text_grob("\u0394 P membership due to bark type (%)       \u0394 P membership due to fire scar (%)", face = "bold", rot = 90, vjust = 0.1, hjust = 0.47), right = text_grob("Resprouting position", face = "bold", rot = 270, hjust = 0.65, vjust = 0.25))

combined_DSF_mature_diff_plot <-  annotate_figure(combined_DSF_mature_diff_plot, top = text_grob("-----------------------------------------------------------------------------------------------------------", vjust = 31.7)) 

ggsave("combined_DSF_mature_diff_plot.png")

combined_DSF_mature_diff_plot
```

```{r fig.height = 8, fig.width = 7}
library(ggpubr)
library(egg)
fig <- egg::ggarrange(WSF.RP.poles.diff.scar, WSF.RP.poles.diff.bark, ncol = 1, nrow = 2) 

combined_WSF_poles_diff_plot <- annotate_figure(fig, bottom = text_grob("Drought severity", face = "bold", vjust = -4.7), top = text_grob("Bark type", face = "bold", vjust = 0.4), left = text_grob("\u0394 P membership due to bark type (%)       \u0394 P membership due to fire scar (%)", face = "bold", rot = 90, vjust = 0.1, hjust = 0.47), right = text_grob("Resprouting position", face = "bold", rot = 270, hjust = 0.65, vjust = 0.25))

combined_WSF_poles_diff_plot <-  annotate_figure(combined_WSF_poles_diff_plot, top = text_grob("-----------------------------------------------------------------------------------------------------------", vjust = 31.7)) 

ggsave("combined_WSF_poles_diff_plot.png")

combined_WSF_poles_diff_plot
```

```{r fig.height = 8, fig.width = 7}
library(ggpubr)
library(egg)
fig <- egg::ggarrange(WSF.RP.mature.diff.scar, WSF.RP.mature.diff.bark, ncol = 1, nrow = 2) 

combined_WSF_mature_diff_plot <- annotate_figure(fig, bottom = text_grob("Drought severity", face = "bold", vjust = -4.7), top = text_grob("Bark type", face = "bold", vjust = 0.4), left = text_grob("\u0394 P membership due to bark type (%)       \u0394 P membership due to fire scar (%)", face = "bold", rot = 90, vjust = 0.1, hjust = 0.47), right = text_grob("Resprouting position", face = "bold", rot = 270, hjust = 0.65, vjust = 0.25))

combined_WSF_mature_diff_plot <-  annotate_figure(combined_WSF_mature_diff_plot, top = text_grob("-----------------------------------------------------------------------------------------------------------", vjust = 31.7)) 

ggsave("combined_WSF_mature_diff_plot.png")

combined_WSF_mature_diff_plot
```
```

